\chapter{Interview Guide}\label{appendix:interview-guide}

The following sections present a semi-structured guideline for the interviews.
The actual conducted interview process may differ from this guide.

\section{Pre-Interview}

\begin{itemize}
	\item Introduction of interviewer
	\item Acknowledgement of interviewee
	\item Pointer to research objective
	\item Indication of anonymous handling of interview data
	\begin{itemize}
		\item Audio recording will transcribed into written form, anonymised, aggregated
		\item Results will be published in master thesis at university
		\item Any personal data will be deleted afterwards
	\end{itemize}
	\item Seeking consent
	\begin{itemize}
		\item Recording of interview for later evaluation
		\item Anonymity is ensured
		\item Inferences about person are not possible
	\end{itemize}
	\item Open questions
	\item Begin recording
\end{itemize}

\section{Person and Organisation}

\begin{itemize}
	\item Age
	\item Current position/role
	\item Years of experience in organisation
	\item Years of experience in current position/role
	\item Core business of organisation
	\item Sector/branch of organisation
\end{itemize}

\section{Definition of terms}

\begin{itemize}
	\item GitOps
	\item Promotion
	\item Release
	\item Environment
\end{itemize}

\section{Questions}\label{appendix:interview-questions}

The following interview questions (IQs) were identified:

\begin{itemize}
	\item IQ 1: What is your understanding of the term GitOps?
	\begin{itemize}
		\item IQ 1.1: Do you follow the OpenGitOps standard?
		\item IQ 1.2: How is GitOps implemented in your organisation?
	\end{itemize}

	\item IQ 2: What problems do you have with promotions in GitOps?
	\begin{itemize}
		\item IQ 2.1: Did you also have these problems before implementing GitOps?
		\item IQ 2.2: How would you imagine a solution?
	\end{itemize}

	\item IQ 3: How do you promote a new release to different environments?
	\begin{itemize}
	        \item IQ 3.1: What is a release in your case?
		\item IQ 3.2: What configuration management tool do you use (e.g. Helm, Kustomize, etc.)?
		\item IQ 3.3: How do your environments look like?
	\end{itemize}
	
	\item IQ 4: Do you practice Continuous Deployment (commit to production is fully automated)?

\end{itemize}

\section{Post-Interview}

\begin{itemize}
	\item End recording
	\item Expression of gratitude for participation
	\item Announcement of results
	\item Contacts for further interviews
\end{itemize}
